% MakeHaven Conflict of Interest Policy

% Needed for TeXShop on Mac OS X
%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\documentclass[12pt]{article}

\input{header.tex}

\title{MakeHaven Conflicts of Interest Policy}
\author{MakeHaven\\a Connecticut non--profit Corporation}

\begin{document}
\maketitle



\section{Purpose}

The purpose of the conflicts of interest policy is to protect the 
Corporation's interest when it is contemplating entering into a 
transaction or arrangement that might benefit the private 
interest of an officer or director of the Corporation. This 
policy is intended to supplement but not replace any applicable 
state laws governing conflicts of interest applicable to 
nonprofit and charitable corporations. 

\section{Definitions}

\subsection{Interested Person}

Any director, principal officer, or member of a committee with 
board delegated powers who has a direct or indirect financial 
interest, as defined below, is an interested person.

\subsection{Financial Interest} 

A person has a financial interest if the person has, directly or 
indirectly, through business, investment or family-- 
\begin{enumerate}
  \item an ownership or investment interest in any entity with 
which the Corporation has a transaction or arrangement; or 
  \item a compensation arrangement with the Corporation or with 
any entity or individual with which the Corporation has a 
transaction or arrangement; or 
  \item a potential ownership or investment interest in, or 
compensation arrangement with, any entity or individual with 
which the Corporation is negotiating a transaction or 
arrangement. 

\end{enumerate}


Compensation includes direct and indirect remuneration as well as 
gifts or favors that are substantial in nature. 

A financial interest is not necessarily a conflict of interest. 
Under~\articlesecref{sec:procedures}{sec:conflict_exists},  a 
person who has a financial interest may have a conflict of 
interest only if the appropriate board or committee decides that 
a conflict of interest exists. 

\section{Procedures}\label{sec:procedures}

\subsection{Duty to Disclose}

In connection with any actual or possible conflicts of interest, 
an interested person must disclose the existence of his or her 
financial interest and must be given the opportunity to disclose 
all material facts to the directors and members of committees 
with board delegated powers considering the proposed transaction 
or arrangement. 

\subsection{Determining Whether a Conflict of Interest Exists}
\label{sec:conflict_exists}

After disclosure of the financial interest and all material 
facts, and after any discussion with the interested person, he/
she shall leave the board or committee meeting while the 
determination of a conflict of interest is discussed and voted 
upon. The remaining board or committee members shall decide if a 
conflict of interest exists. 

\subsection{Procedures for Addressing the Conflict of Interest}

\begin{enumerate}
\item An interested person may make a presentation at the board 
or committee meeting, but after such presentation, he/she shall 
leave the meeting during the discussion of, and the vote on, the 
transaction or arrangement that results in the conflict of 
interest. 

\item The chairperson of the board or committee shall, if 
appropriate, appoint a disinterested person or committee to 
investigate alternatives to the proposed transaction or 
arrangement. 

\item After exercising due diligence, the board or committee 
shall determine whether the Corporation can obtain a more 
advantageous transaction or arrangement with reasonable efforts 
from a person or entity that would not give rise to a conflict of 
interest. 

\item If a more advantageous transaction or arrangement is not 
reasonably attainable under circumstances that would not give 
rise to a conflict of interest, the board or committee shall 
determine by a majority vote of the disinterested directors 
whether the transaction or arrangement is in the Corporation's 
best interest and for its own benefit and whether the transaction 
is fair and reasonable to the Corporation and shall make its 
decision as to whether to enter into the transaction or 
arrangement in conformity with such determination. 
\end{enumerate}

\subsection{Violations of the Conflicts of Interest Policy}

\begin{enumerate}
\item If the board or committee has reasonable cause to believe 
that a member has failed to disclose actual or possible conflicts 
of interest, it shall inform the member of the basis for such 
belief and afford the member an opportunity to explain the 
alleged failure to disclose. 

\item If, after hearing the response of the member and making 
such further investigation as may be warranted in the 
circumstances, the board or committee determines that the member 
has in fact failed to disclose an actual or possible conflict of 
interest, it shall take appropriate disciplinary and corrective 
action. 
\end{enumerate}

\section{Records of Proceedings}

The minutes of the board and all committees with board-delegated 
powers shall contain---

\begin{enumerate}

\item the names of the persons who disclosed or otherwise were 
found to have a financial interest in connection with an actual 
or possible conflict of interest, the nature of the financial 
interest, any action taken to determine whether a conflict of 
interest was present, and the board's or committee's decision as 
to whether a conflict of interest in fact existed. 

\item the names of the persons who were present for discussions 
and votes relating to the transaction or arrangement, the content 
of the discussion, including any alternatives to the proposed 
transaction or arrangement, and a record of any votes taken in 
connection therewith. 
\end{enumerate}

\section{Compensation}

\begin{enumerate}
\item A voting member of the board of directors who receives 
compensation, directly or indirectly, from the Corporation for 
services is precluded from voting on matters pertaining to that 
member's compensation. 

\item An individual who is a voting member of the board of 
directors and receives compensation, directly or indirectly, from 
the Corporation for services is precluded from discussing and 
voting on matters pertaining to that member's and other members' 
compensation. No person, either individually or collectively, is 
prohibited from providing information to the board of directors 
regarding compensation. 

\item A voting member of any committee whose jurisdiction 
includes compensation matters and who receives compensation, 
directly or indirectly, from the Corporation for services is 
precluded from voting on matters pertaining to that member's 
compensation. 

\item Members who receive compensation, directly or indirectly, 
from the Corporation, whether as employees or independent 
contractors, are precluded from membership on any committee whose 
jurisdiction includes compensation matters. No member, either 
individually or collectively, is prohibited from providing 
information to any committee regarding compensation. 
\end{enumerate}


\section{Statements}

Each director, principal officer and member of a committee with 
board delegated powers shall annually sign a statement which 
affirms that such person---

\begin{enumerate}
\item has received a copy of the conflicts of interest policy, 

\item has read and understands the policy, 

\item has agreed to comply with the policy, and 

\item understands that the Corporation is a charitable 
organization and that in order to maintain its federal tax 
exemption it must engage primarily in activities which accomplish 
one or more of its tax--exempt purposes. 
\end{enumerate}

\section{Periodic Reviews}\label{sec:reviews}

To ensure that the Corporation operates in a manner consistent 
with its charitable purposes and that it does not engage in 
activities that could jeopardize its status as an organization 
exempt from federal income tax, periodic reviews shall be 
conducted. The periodic reviews shall, at a minimum, include the 
following subjects: 

\begin{enumerate}
\item Whether compensation arrangements and benefits are 
reasonable and are the result of arm's-length bargaining. 

\item Whether acquisitions of equipment and services result in 
inurement or impermissible private benefit. 

\item Whether partnership and joint venture arrangements and 
arrangements with other organizations conform to written 
policies, are properly recorded, reflect reasonable payments for 
goods and services, further the Corporation's charitable purposes 
and do not result in inurement or impermissible private benefit. 

\end{enumerate}

\section{Use of Outside Experts}\label{sec:experts}

In conducting the periodic reviews provided for in Article~
\vref{sec:reviews}, the Corporation may, but need not, use 
outside advisors. If outside experts are used their use shall not 
relieve the board of its responsibility for ensuring that 
periodic reviews are conducted. 

\end{document}  